### 2019-06-20 Datelife_paper1.Rmd
tol plan and poster figs

Information from source chronograms can be summarised with a summary statistic of tree branch
lengths (such as median or mean) or using the Super Distance Matrix (SDM) approach for supertree reconstruction [@Criscuolo2006]. <!-- and the hierarchical clustering method of ttol-->
The resulting summary patristic distance matrix could be clustered with classic algorithms. However, we noticed that the resulting trees are often non-ultrameric and do not reflect the source chronogram data (see datelife_examples package). Instead, we obtained a distribution of age data from the summary matrix available for nodes on a consensus tree. The Branch Length Adjuster (BLADJ) algorithm [@Webb2008] was then used to distribute node ages evenly over the consensus tree, minimizing age variance in the resulting chronogram.

>>>>

Information from source chronograms can be summarised with a summary statistic of tree branch
lengths (such as median or mean) or using the Super Distance Matrix (SDM) approach for supertree reconstruction [@Criscuolo2006]. <!-- and the hierarchical clustering method of ttol-->
The resulting summary patristic distance matrix could be clustered with classic algorithms. However, we noticed that the resulting trees are often non-ultrameric and do not reflect the source chronogram data (see datelife_examples package). Instead, we obtained a distribution of age data from the summary matrix available for nodes on a consensus tree. The Branch Length Adjuster (BLADJ) algorithm [@
2008] was then used to distribute node ages evenly over the consensus tree, minimizing age variance in the resulting chronogram.

### 2019-06-20 Intro.Rmd
tol plan and poster figs

<!--INTRODUCTION: 3rd paragraph: problems associated-->
The TimeTree project [@Hedges2006; @Hedges2015; @Kumar2017] has aggregated chronograms
from 3,163 studies, encompassing 97,085 species [@Kumar2017], and continues to grow.
However, even in this gold standard resource, the included taxa only encompass between
0.097 and 3.236% of total species diversity (following taxonomic expert opinion
on the global, extant species numbers, which ranges from 3 to 100 million species
[Mayr2010; Moran2011]). One advantage of TimeTree is that it includes taxa from
across the tree of life, versus more specialized chronogram repositories focusing on land plants
[@stevens2001apw] and mammals [@webb2005phylomatic], birds [@Jetz2012], and fish [@chang2019r]. Users can choose
between a web interface or a mobile app to receive information on divergence times
for the evolutionary history of a lineage, pairs of taxa, all lineages within a
taxon, or a list of taxa. As a science communication tool, TimeTree project is
very powerful: it has a friendly graphical interface, with informative and colorful
outputs, that allows the general public to satisfy curiosity regarding a particular
organism of interest or group of them. It is of limited utility for scientific
studies, however. The thousands of trees that have been entered are unavailable
for examination or reuse; according to the creators (see TimeTree web FAQ), methods
for allowing data downloading have been under discussion for the past several years
yet the primary data remain closed. Moreover, there is no Application Programming
Interface (API) allowing programmatic access to any data, greatly impairing the
possibility of large-scale, automated data-mining, which is not allowed under TimeTree
website's terms of use. The nearly hundred thousand taxon summary chronogram generated
from TimeTree resources is not available with its publication [@Kumar2017] or the
TimeTree website, though the still substantial chronogram from a previous publication
[@Hedges2015] was made available at OToL.
<!--What led to datelife development?
Describe public and research necessities covered by datelife
It is important to use available data on time frame of lineage origin:
To know the state of dating for a group of interest:
What range of estimated ages exist already? Are fossil and molecular
time frames coherent? (e.g., [@Magallon2015c]).
To construct a time tree of life.
For science communication, improve scientific discussions, time-framing
other events of importance in other research areas.
Comparative method: autocorrelation (phylogenetic relatedeness) corrections, phylogenetic signal
-->

>>>>

<!--INTRODUCTION: 3rd paragraph: problems associated-->
The TimeTree project [@Hedges2006; @Hedges2015; @Kumar2017] has aggregated chronograms
from 3,163 studies, encompassing 97,085 species [@Kumar2017], and continues to grow.
However, even in this gold standard resource, the included taxa only encompass between
0.097 and 3.236% of total species diversity (following taxonomic expert opinion
on the global, extant species numbers, which ranges from 3 to 100 million species
[@Mayr2010; @Mora2011]). One advantage of TimeTree is that it includes taxa from
across the tree of life, versus more specialized chronogram repositories focusing on land plants
[@stevens2001apw] and mammals [@webb2005phylomatic], birds [@Jetz2012], and fish [@chang2019r]. Users can choose
between a web interface or a mobile app to receive information on divergence times
for the evolutionary history of a lineage, pairs of taxa, all lineages within a
taxon, or a list of taxa. As a science communication tool, TimeTree project is
very powerful: it has a friendly graphical interface, with informative and colorful
outputs, that allows the general public to satisfy curiosity regarding a particular
organism of interest or group of them. It is of limited utility for scientific
studies, however. The thousands of trees that have been entered are unavailable
for examination or reuse; according to the creators (see TimeTree web FAQ), methods
for allowing data downloading have been under discussion for the past several years
yet the primary data remain closed. Moreover, there is no Application Programming
Interface (API) allowing programmatic access to any data, greatly impairing the
possibility of large-scale, automated data-mining, which is not allowed under TimeTree
website's terms of use. The nearly hundred thousand taxon summary chronogram generated
from TimeTree resources is not available with its publication [@Kumar2017] or the
TimeTree website, though the still substantial chronogram from a previous publication
[@Hedges2015] was made available at OToL.
<!--What led to datelife development?
Describe public and research necessities covered by datelife
It is important to use available data on time frame of lineage origin:
To know the state of dating for a group of interest:
What range of estimated ages exist already? Are fossil and molecular
time frames coherent? (e.g., [@Magallon2015c]).
To construct a time tree of life.
For science communication, improve scientific discussions, time-framing
other events of importance in other research areas.
Comparative method: autocorrelation (phylogenetic relatedeness) corrections, phylogenetic signal
-->
